Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-06-01 Documents/M1/dgM1Curvilinear.tex
Updates to manuscript.

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\UDG}{\bU_{\mbox{\tiny DG}}}
\newcommand{\sumx}{\sum_{i=1}^{d}}

@@

For general curvilinear spatial coordinates $\{x^{i}\}$, the two-moment model for radiation transport takes the form \citep[see, e.g.,][for fully general relativistic treatments]{shibata_etal_2011,cardall_etal_2013a}
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\cJ\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\cH^{i}\,\Big)=\chi\,\big(\cJ_{0}-\cJ\big),
  \label{eq:momentEquationEnergy}
\end{equation}
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\cH_{j}\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\cK^{i}_{~j}\,\Big)
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\kappa\,\cH_{j},
  \label{eq:momentEquationMomentum}
\end{equation}
where $\cJ$, $\cH^{i}$, and $\cK^{ij}$ are defined as angular moments of a positive neutrino distribution function $f(\omegaNu,\epsilonNu,\bx,t)$
\begin{equation}
  \big\{\,\cJ,\,\cH^{i},\,\cK^{ij}\,\big\}(\epsilonNu,\bx,t)
  =\int_{\bbS^{2}}f(\omegaNu,\epsilonNu,\bx,t)\,\big\{1,l^{i},l^{i}\,l^{j}\}\,d\omegaNu.  
  \label{eq:angularMoments}
\end{equation}
Here, $\{l^{i}\}$ are components of a unit three-vector proportional to the neutrino three-momentum $\bp=\epsilonNu\,\bl$.  
(We adopt the Einstein summation convention and let repeated Latin indices run from $1$ to $3$.)
In a kinetic description, neutrinos are described by the phase space distribution function $f$, and the momentum space coordinates used here are the neutrino energy $\epsilonNu$, and the angles $\thetaNu$ and $\phiNu$ (describing the neutrino propagation direction relative to a local orthonormal basis).  
In forming the angular moments in Eq.~\eqref{eq:angularMoments}, the integrals extend over the unit sphere
\begin{equation}
  \bbS^{2}=\big\{\,(\thetaNu,\phiNu)~|~\thetaNu\in[0,\pi),\,\phiNu\in[0,2\pi)\,\big\}.  
\end{equation}
The angular moments are related to the neutrino energy density, momentum density, and stress ($\cJ$, $\cH^{i}$, and $\cK^{ij}$, respectively).  
They are functions of the neutrino energy $\epsilonNu$, spatial position $\{x^{i}\}$, and time $t$.

To model neutrino matter interactions, we include emission and absorption, and isotropic, elastic scattering on nucleons and nuclei.  
On the right-hand sides of Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, $\chi(\epsilonNu,\bU)$ and $\kappa(\epsilonNu,\bU)=\chi(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$ are absorption and total (absorption plus scattering) opacities, respectively.  
The opacities depend on the neutrino energy and the local thermodynamic state of the fluid $\bU=(\rho,T,Y_{e})^{T}$, where $\rho$ is the (baryon) mass density, $T$ the temperature, and $Y_{e}$ the electron fraction.  
We use the opacities given by \citet{bruenn_1985}.  
In Eq.~\eqref{eq:momentEquationEnergy}, the equilibrium density $\cJ_{0}=4\pi\,f_{0}(\epsilonNu,\bU)$ is computed from the angular moment of the isotropic Fermi-Dirac distribution
\begin{equation}
  f_{0}(\epsilonNu,\bU)=\f{1}{e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}+1},
\end{equation}
where $k_{\mbox{\tiny B}}$ is Boltzmann's constant, and $\mu_{\nu}$ is the neutrino chemical potential.

@@

We define the geometry sources in Eq.~\eqref{eq:momentEquationMomentum} by
\begin{equation}
  \cG_{j}
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric} and writing in terms of $\cK^{i}_{~j}$, the components are
\begin{equation}
  \cG_{1}
  =\f{1}{2}\,\Big(\,\cK^{2}_{~2}\,\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}+\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}\,\Big), \quad
  \cG_{2}
  =\f{1}{2}\,\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}, \quad\text{and}\quad
  \cG_{3}
  =0.
\end{equation}
Then, by defining the vector of evolved moments $\bcM=\big(\,\cJ,\,\cH_{1},\,\cH_{2},\,\cH_{3}\,\big)^{T}$, the vector of geometry sources $\bcG(\bcM)=\big(\,0,\,\cG_{1},\,\cG_{2},\,\cG_{3}\,\big)^{T}$, the vector of collision sources $\bcC(\bcM)=\big(\,\chi\,(\cJ_{0}-\cJ),\,-\kappa\,\cH_{1},\,-\kappa\,\cH_{2},\,-\kappa\,\cH_{3}\,\big)^{T}$, and the flux vectors $\bcF^{i}(\bcM)=\big(\,\cH^{i},\,\cK_{~1}^{i},\,\cK_{~2}^{i},\,\cK_{~3}^{i}\,\big)^{T}$, we write the system of equations in compact form
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\bcM\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\bcF^{i}(\bcM)\,\Big)
  =\bcG(\bcM)+\bcC(\bcM).
  \label{eq:momentEquationsCompact}
\end{equation}
Eq.~\eqref{eq:momentEquationsCompact} forms the basis for developing DG methods for neutrino transport.

@@

\section{Neutrino Opacities}

@@

The last two terms are contributions from isoenergetic scattering of neutrino by nucleons and nuclei. The isoenergetic scattering kernel $R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)$ has the following relation:
\begin{align}
R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) & = R_{IS}^{nucleons}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) +  R_{IS}^{nuclei}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)\\
& = \int_{0}^{2\pi}d
\end{align}
\clearpage

@@

To measure the accuracy of the implementation using Cartesian coordinates, and to gauge the performance of the DG scheme using different limiter parameters, we advect a sine wave across the one-dimensional, periodic domain $D=\{x:x\in[0,1]\}$ ten times, until $t=10$.  
We let the initial condition be given by
\begin{equation}
  \cJ(x,t=0)=\cH_{x}(x,t=0)=1+\sin\big(2\pi\,x\big).  
\end{equation}
Results are shown in Figure~\ref{fig:streamingSineWave} and Table~\ref{tab:streamingSineWave}.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.38]{./Figures/StreamingSineWave1D}
  \end{center}
  \caption{Energy density versus position after ten grid crossings ($t=10$) for the streaming sine wave problem computed with the DG(2)+RK3 scheme using 16 elements.  We compare results obtained using various limiter parameters (see text for details) with the analytic solution (dotted line).}
  \label{fig:streamingSineWave}
\end{figure}

@@

Snapshots of the solution ($\cJ$ versus $r$), obtained with the DG(2)+RK3 scheme, are shown in Figure~\ref{fig:sphericalWave} for $t=\{2.5,3.5,5.0,7.0\}$ (open circles).  
For comparison, the analytical solution is plotted with dotted lines.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.38]{./Figures/GaussianSphericalWave1D}
  \end{center}
  \caption{Energy density versus radius at various times for the spherical wave problem computed with the DG(2)+RK3 scheme using 32 elements.  The numerical solution (open circles) is compared to the analytical solution (dotted lines).  }
  \label{fig:sphericalWave}
\end{figure}

\begin{table}
  \begin{center}
  \caption{$L^{\infty}$ error and convergence rates for the spherical wave problem.}
  \begin{tabular}{ccccccccc}
    & \multicolumn{4}{c}{SSP-RK2} & \multicolumn{4}{c}{SSP-RK3} \\
    \cmidrule(r){2-5} \cmidrule(r){6-9}
    $N$ & DG(1) & Rate & DG(2) & Rate & DG(2) & Rate & DG(3) & Rate \\
    \midrule \midrule
    32   & $4.648\times10^{-4}$ & ---  & $5.540\times10^{-5}$ & ---   & $4.803\times10^{-5}$ & ---  & $1.707\times10^{-5}$ & --- \\
    64   & $9.506\times10^{-5}$ &2.29& $1.088\times10^{-5}$ &2.35& $1.373\times10^{-5}$ &1.81& $7.812\times10^{-7}$ &4.45 \\
    128 & $1.996\times10^{-5}$ &2.25& $1.502\times10^{-6}$ &2.86& $1.836\times10^{-6}$ &2.90& $1.150\times10^{-7}$ &2.76 \\
    256 & $5.783\times10^{-6}$ &1.79& $2.529\times10^{-7}$ &2.57& $1.895\times10^{-7}$ &3.28& $6.268\times10^{-9}$ &4.20 \\
    \midrule \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

To test the ability of the DG method to capture discontinuities, we solve the Riemann problem presented by \citet{OHF2012}.  
To compare with their results, we use the Eddington factor due to Levermore \citep{levermore_1984}
\begin{equation}
  \psi(h)=\f{1}{3}\big(\,5-2\,\sqrt{4-3\,h^{2}}\,\big).  
  \label{eq:eddingtonFactorLevermore}
\end{equation}
The computational domain extends from $x=-0.05$ to $x=0.1$, and a discontinuity is located at $x_{\mbox{\tiny d}}=0.0$.  
The initial condition is given by
\begin{equation}
  \bcM(x,t=0)=
  \left\{
  \begin{array}{l}
  \bcM_{\mbox{\tiny L}}=\big(1,0.9999,0,0\big)^{T} \quad\text{if}\quad x\le x_{\mbox{\tiny d}} \\
  \bcM_{\mbox{\tiny R}}=\big(0.5,0,0,0\big)^{T} \quad\text{otherwise}.
  \end{array}
  \right.
  \label{eq:riemannProblem}
\end{equation}
Results for $t=0.1$ are shown in Figure~\ref{fig:riemannProblem}.  
The discontinuities remain sharp, and oscillations around the discontinuities are controlled (in part due to limiting of characteristic variables; cf. Section~\ref{sec:slopeLimiters}).

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.4]{./Figures/RiemannProblem1D} &
      \includegraphics[scale=0.4]{./Figures/RiemannProblem1D_Limiters}
    \end{tabular}
  \end{center}
  \caption{Results from the one-dimensional Riemann problem in Eq.~\eqref{eq:riemannProblem}, obtained with the DG(2)+RK3 method using $240$ elements, the HLL Riemann solver, and the Eddington factor in Eq.~\eqref{eq:eddingtonFactorLevermore}.  
  The number density $\cJ$ (black) and the number flux density $\cH_{r}$ (red) are plotted for $t=0.1$.}
  \label{fig:riemannProblem}
\end{figure}

\subsection{Diffusion Problem}

This problem is adapted from \citet{abdikamalov_etal_2012} \citep[see also][]{pons_etal_2000,sumiyoshiYamada_2012}.

\begin{equation}
  \cJ(r,t)=\Big(\f{t_{0}}{t_{0}+t}\Big)^{3/2}\,\exp\Big\{\,-\f{3\,\sigma\,r^{2}}{4\,(t_{0}+t)}\,\Big\}
\end{equation}

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-5_Density} &
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-5_Flux}
    \end{tabular}
  \end{center}
  \caption{Results from the one-dimensional diffusion problem in spherical symmetry with $\sigma=10^{-5}$~cm$^{-1}$, obtained with the DG(2)+SIRK2 method using $24$ elements.  
  The number density $\cJ$ (left panel) and the number flux density $\cH_{r}$ (right panel) are plotted for $t=0,0.3,1.0$~ms (red, blue, and black curves, respectively).  
  Solid lines represent the analytical solution while open circles represent the element center value of the numerical solution.}
  \label{fig:diffusionProblem1e-5}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-1_Density} &
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-1_Flux}
    \end{tabular}
  \end{center}
  \caption{Results similar to those displayed in Figure~\ref{fig:diffusionProblem1e-5}, but with $\sigma=10^{-1}$~cm$^{-1}$.}
  \label{fig:diffusionProblem1e-1}
\end{figure}

@@

>>>>

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\UDG}{\bU_{\mbox{\tiny DG}}}
\newcommand{\sumx}{\sum_{i=1}^{d}}
\newcommand{\NES}{\mbox{\tiny NES}}

@@

For general curvilinear spatial coordinates $\{x^{i}\}$, the two-moment model for radiation transport takes the form \citep[see, e.g.,][for fully general relativistic treatments]{shibata_etal_2011,cardall_etal_2013a}
\begin{align}
  &\f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\cJ\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\cH^{i}\,\Big)
  =\tilde{\chi}\,\big(\cJ_{0}-\cJ\big) \nonumber \\
  &\hspace{12pt}
  +\big(4\pi-\cJ\big)\int_{\bbR^{+}}R^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\cJ\int_{\bbR^{+}}R^{\Out}(\epsilonNu,\epsilonNu')\,\big(4\pi-\cJ(\epsilonNu')\big)\,dV_{\epsilonNu'},
  \label{eq:momentEquationEnergy}
\end{align}
\begin{align}
  &\f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\cH_{j}\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\cK^{i}_{~j}\,\Big)
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\kappa\,\cH_{j} \nonumber \\
  &\hspace{12pt}
  -\cH_{j}\int_{\bbR^{+}}\big[\,R^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')+R^{\Out}(\epsilonNu,\epsilonNu')\big(4\pi-\cJ(\epsilonNu')\big)\,\big]\,dV_{\epsilonNu'},
  \label{eq:momentEquationMomentum}
\end{align}
where $\cJ$, $\cH^{i}$, and $\cK^{ij}$ are defined as angular moments of a positive neutrino distribution function $f(\omegaNu,\epsilonNu,\bx,t)$
\begin{equation}
  \big\{\,\cJ,\,\cH^{i},\,\cK^{ij}\,\big\}(\epsilonNu,\bx,t)
  =\int_{\bbS^{2}}f(\omegaNu,\epsilonNu,\bx,t)\,\big\{1,l^{i},l^{i}\,l^{j}\}\,d\omegaNu.  
  \label{eq:angularMoments}
\end{equation}
Here, $\{l^{i}\}$ are components of a unit three-vector proportional to the neutrino three-momentum $\bp=\epsilonNu\,\bl$.  
(We adopt the Einstein summation convention and let repeated Latin indices run from $1$ to $3$.)
In a kinetic description, neutrinos are described by the phase space distribution function $f$, and the momentum space coordinates used here are the neutrino energy $\epsilonNu$, and the angles $\thetaNu$ and $\phiNu$ (describing the neutrino propagation direction relative to a local orthonormal basis).  
In forming the angular moments in Eq.~\eqref{eq:angularMoments}, the integrals extend over the unit sphere
\begin{equation}
  \bbS^{2}=\big\{\,(\thetaNu,\phiNu)~|~\thetaNu\in[0,\pi),\,\phiNu\in[0,2\pi)\,\big\}.  
\end{equation}
The angular moments are related to the neutrino energy density, momentum density, and stress ($\cJ$, $\cH^{i}$, and $\cK^{ij}$, respectively).  
They are functions of the neutrino energy $\epsilonNu$, spatial position $\{x^{i}\}$, and time $t$.

To model neutrino matter interactions, we include emission and absorption, (isotropic) elastic scattering on nucleons and nuclei, and neutrino-electron scattering (NES).  
On the right-hand sides of Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, $\tilde{\chi}(\epsilonNu,\bU)$ and $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$ are absorption and total (absorption plus scattering) opacities, respectively.  
The absorption opacity $\tilde{\chi}$ is corrected for stimulated absorption \citep[cf.][]{mezzacappaBruenn_1993a}.  
The NES rates $R^{\In}$ and $R^{\Out}$ are obtained by expanding the full kernels in a Legendre series, retaining only the isotropic part \citep[see, e.g.,][]{cernohorsky_1994}, and the integrals extend over all energies $\bbR^{+}=\{\epsilonNu~|~\epsilonNu\in[0,\infty)\}$.  
The energy volume element is $dV_{\epsilonNu}=\epsilonNu^{2}d\epsilonNu$.  
The opacities depend on the neutrino energy and the local thermodynamic state of the fluid $\bU=(\rho,T,Y_{e})^{T}$, where $\rho$ is the (baryon) mass density, $T$ the temperature, and $Y_{e}$ the electron fraction.  
For simplicity we consider a single neutrino species (electron neutrinos), and we use opacities given by \citet{bruenn_1985}; see Section~\ref{sec:opacities} for further details.  
In Eq.~\eqref{eq:momentEquationEnergy}, the equilibrium density $\cJ_{0}=4\pi\,f_{0}(\epsilonNu,\bU)$ is computed from the angular moment of the isotropic Fermi-Dirac distribution
\begin{equation}
  f_{0}(\epsilonNu,\bU)=\f{1}{e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}+1},
\end{equation}
where $k_{\mbox{\tiny B}}$ is Boltzmann's constant, and $\mu_{\nu}=\mu_{p}-\mu_{n}+\mu_{e}$ is the neutrino chemical potential.  
(The chemical potentials of protons, neutrons, and electrons are denoted by $\mu_{p}$, $\mu_{n}$, and $\mu_{e}$, respectively.)

@@

For the forthcoming description it is useful to write the moment equations in a more compact form.  
To this end, we define the geometry sources in Eq.~\eqref{eq:momentEquationMomentum} by
\begin{equation}
  \cG_{j}
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric} and writing in terms of $\cK^{i}_{~j}$, the components are
\begin{equation}
  \cG_{1}
  =\f{1}{2}\,\Big(\,\cK^{2}_{~2}\,\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}+\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}\,\Big), \quad
  \cG_{2}
  =\f{1}{2}\,\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}, \quad\text{and}\quad
  \cG_{3}
  =0.
\end{equation}
(Explicit expressions for metric functions appearing in the geometry sources are listed in Table~\ref{tab:metricFunctions}.)  
Then, by defining the vector of evolved moments $\bcM=\big(\,\cJ,\,\bcH\,\big)^{T}$, the vector of geometry sources $\bcG(\bcM)=\big(\,0,\,\cG_{1},\,\cG_{2},\,\cG_{3}\,\big)^{T}$, the vector of collision sources $\bcC(\bcM)=\big(\,\chi\,(\cJ_{0}-\cJ)+\cL_{\NES},\,-(\kappa+\kappa_{\NES})\,\bcH\,\big)^{T}$, and the flux vectors $\bcF^{i}(\bcM)=\big(\,\cH^{i},\,\cK_{~1}^{i},\,\cK_{~2}^{i},\,\cK_{~3}^{i}\,\big)^{T}$, we write the system of equations in compact form
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\bcM\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\bcF^{i}(\bcM)\,\Big)
  =\bcG(\bcM)+\bcC(\bcM).
  \label{eq:momentEquationsCompact}
\end{equation}
Here, $\cL_{\NES}$ and $\kappa_{\NES}$ represent the integral operators appearing in Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, respectively.  
Eq.~\eqref{eq:momentEquationsCompact} forms the basis for developing DG methods for neutrino transport.

@@

\section{Neutrino Opacities}
\label{sec:opacities}

@@

The last two terms are contributions from isoenergetic scattering of neutrino by nucleons and nuclei. The isoenergetic scattering kernel $R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)$ has the following relation:
\begin{align}
R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) & = R_{IS}^{nucleons}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) +  R_{IS}^{nuclei}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)\\
& = \int_{0}^{2\pi}d
\end{align}

@@

To measure the accuracy of the implementation using Cartesian coordinates, and to gauge the performance of the DG scheme using different limiter parameters, we advect a sine wave across the one-dimensional, periodic domain $D=\{x:x\in[0,1]\}$ ten times, until $t=10$.  
We let the initial condition be given by
\begin{equation}
  \cJ(x,t=0)=\cH_{x}(x,t=0)=1+\sin\big(2\pi\,x\big).  
\end{equation}
Results are shown in Figure~\ref{fig:streamingSineWave} and Table~\ref{tab:streamingSineWave}.  
In Figure~\ref{fig:streamingSineWave}, results obtained using the TVD limiter (cf. Eq.~\eqref{eq:minmodB} with $M=0$) with various $\beta$ values are shown.  
With this limiter, the accuracy of the scheme is severely compromised.  
The accuracy can be significantly improved with the TVB limiter ($M>0$), but the proper value for $M$ is problem dependent, and difficult to set a priori.  
Instead, we have implemented the discontinuity detector based on Harten's subcell resolution method as a ``troubled-cell'' indicator \citep{harten_1989,qiuShu_2005}, which prevents limiting in smooth regions.  
Results using the detector (black open circles) show a significant improvement.  
In Table~\ref{tab:streamingSineWave} we list convergence results for this latter method, which demonstrates high-order accuracy (i.e., better than second-order) when expected.  
Unless otherwise noted, we continue to use the discontinuity detector in the remaining numerical experiments.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.78]{./Figures/StreamingSineWave1D}
  \end{center}
  \caption{Particle density versus position after ten grid crossings ($t=10$) for the streaming sine wave problem computed with the DG(2)+RK3 scheme using 16 elements.  
  We compare results obtained using various limiter parameters (open circles; see text for details) with the analytic solution (dotted line).  
  Results obtained using the discontinuity detector (D) is denoted with black open circles.}
  \label{fig:streamingSineWave}
\end{figure}

@@

Snapshots of the solution ($\cJ$ versus $r$), obtained with the DG(2)+RK3 scheme, are shown in Figure~\ref{fig:sphericalWave} for $t=\{2.5,3.5,5.0,7.0\}$ (open circles).  
For comparison, the analytical solution is plotted with dotted lines.  
There is good agreement between the numerical and analytical solutions.  
In Table~\ref{tab:sphericalWave} we list convergence results obtained with the various schemes.  
High-order accuracy is achieved with the DG(2)+RK3 and DG(3)+RK3 schemes.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.78]{./Figures/GaussianSphericalWave1D}
  \end{center}
  \caption{Energy density versus radius at various times for the spherical wave problem computed with the DG(2)+RK3 scheme using 32 elements.  The numerical solution (open circles) is compared to the analytical solution (dotted lines).  }
  \label{fig:sphericalWave}
\end{figure}

\begin{table}
  \begin{center}
  \caption{$L^{\infty}$ error and convergence rates for the spherical wave problem.}
  \label{tab:sphericalWave}
  \begin{tabular}{ccccccccc}
    & \multicolumn{4}{c}{SSP-RK2} & \multicolumn{4}{c}{SSP-RK3} \\
    \cmidrule(r){2-5} \cmidrule(r){6-9}
    $N$ & DG(1) & Rate & DG(2) & Rate & DG(2) & Rate & DG(3) & Rate \\
    \midrule \midrule
    32   & $4.648\times10^{-4}$ & ---  & $5.540\times10^{-5}$ & ---   & $4.803\times10^{-5}$ & ---  & $1.707\times10^{-5}$ & --- \\
    64   & $9.506\times10^{-5}$ &2.29& $1.088\times10^{-5}$ &2.35& $1.373\times10^{-5}$ &1.81& $7.812\times10^{-7}$ &4.45 \\
    128 & $1.996\times10^{-5}$ &2.25& $1.502\times10^{-6}$ &2.86& $1.836\times10^{-6}$ &2.90& $1.150\times10^{-7}$ &2.76 \\
    256 & $5.783\times10^{-6}$ &1.79& $2.529\times10^{-7}$ &2.57& $1.895\times10^{-7}$ &3.28& $6.268\times10^{-9}$ &4.20 \\
    \midrule \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

To test the ability of the DG method to capture discontinuities without spurious oscillations, we solve the Riemann problem presented by \citet{OHF2012}.  
This problem is also challenging because the condition $\cJ-|\bcH|>0$ can be violated.  
To compare with the results of \citet{OHF2012}, we use the Eddington factor due to Levermore \citep{levermore_1984}
\begin{equation}
  \psi(h)=\f{1}{3}\big(\,5-2\,\sqrt{4-3\,h^{2}}\,\big).  
  \label{eq:eddingtonFactorLevermore}
\end{equation}
The computational domain extends from $x=-0.05$ to $x=0.1$, and a discontinuity is located at $x_{\mbox{\tiny d}}=0.0$.  
The initial condition is given by
\begin{equation}
  \bcM(x,t=0)=
  \left\{
  \begin{array}{l}
  \bcM_{\mbox{\tiny L}}=\big(1,0.9999,0,0\big)^{T} \quad\text{if}\quad x\le x_{\mbox{\tiny d}} \\
  \bcM_{\mbox{\tiny R}}=\big(0.5,0,0,0\big)^{T} \quad\text{otherwise}.
  \end{array}
  \right.
  \label{eq:riemannProblem}
\end{equation}
Results for $t=0.1$ are shown in Figure~\ref{fig:riemannProblem}.  
In the left panel we plot the number density and flux, obtained with the third-order method (DG(2)+RK3).  
The discontinuities remain sharp, and oscillations around the discontinuities are controlled (in part due to limiting of characteristic variables; cf. Section~\ref{sec:slopeLimiters}).  
In the right panel we plot the number density for three computations: limiting triggered everywhere (blue), no slope limiting (positivity limiting only; red), and limiting triggered by the discontinuity detector (black).  
In the plot we have zoomed in on the discontinuity around $x=0.075$.  
Without limiting, significant oscillations are present in the solution.  
With the slope limiter, the oscillations are essentially removed, while using the discontinuity detector to trigger limiting maintains a sharper discontinuity than when limiting is permitted everywhere.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.4]{./Figures/RiemannProblem1D} &
      \includegraphics[scale=0.4]{./Figures/RiemannProblem1D_Limiters}
    \end{tabular}
  \end{center}
  \caption{Results from the one-dimensional Riemann problem in Eq.~\eqref{eq:riemannProblem}, obtained with the DG(2)+RK3 method using $240$ elements, the HLL Riemann solver, and the Eddington factor in Eq.~\eqref{eq:eddingtonFactorLevermore}.  
  The number density $\cJ$ (black) and the number flux density $\cH_{1}$ (red) are plotted for $t=0.1$.}
  \label{fig:riemannProblem}
\end{figure}

\subsection{Spherical Diffusion}

Next, we consider a diffusion problem in spherical symmetry, which involves a constant scattering opacity $\sigma$.  
This problem is adapted from \citet{abdikamalov_etal_2012} \citep[see also][]{pons_etal_2000,sumiyoshiYamada_2012}.  
For sufficiently high scattering opacity, the moment equations limit to a diffusion equation for the radiation number density.  
With a Gaussian initial distribution for the energy density $\cJ_{0}(x)=\exp(-3\,\sigma\,r^{2}/(4\,t_{0}))$, the analytical solution to the limiting diffusion equation is given by
\begin{equation}
  \cJ(r,t)=\Big(\f{t_{0}}{t_{0}+t}\Big)^{3/2}\,\exp\Big\{\,-\f{3\,\sigma\,r^{2}}{4\,(t_{0}+t)}\,\Big\},
\end{equation}
while the number flux is obtained from $\cH_{r}=-\pd{\cJ}{r}/(3\,\sigma)$.

We present two versions of this test:  one with $\sigma=10^{-5}$~cm$^{-1}$ (thin test), and one with $\sigma=10^{-1}$~cm$^{-1}$ (thick test).  
In the thin test, the computational domain extends over $r\in[0,100]$~km, $t_{0}=0.3$~ms, and is run until $t=1$~ms.  
In the thick test, the computational domain extends over $r\in[0,50]$~km and $t_{0}=0.3$~s, and is run until $t=1$~s.  
In both tests we use 32 elements and the DG(2)+SIRK2 scheme.  
With the mean-free path defined as $\lambda=1/\sigma$, the ratio of the mean-free path to the cell width $\Delta r$ is $0.32$ in the thin test, and $6.4\times10^{-5}$ in the thick test.

Results are plotted in Figure~\ref{fig:diffusionProblem}.  
There is good agreement between the numerical and analytical solutions, for both the thin and thick cases.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-5_Density} &
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-5_Flux} \\
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-1_Density} &
      \includegraphics[scale=0.38]{./Figures/GaussianSphericalDiffusion_Kappa_1e-1_Flux}
    \end{tabular}
  \end{center}
  \caption{Results from the one-dimensional diffusion problem in spherical symmetry, obtained with the DG(2)+SIRK2 method using $32$ elements.  
  The number density $\cJ$ (left panel) and the number flux density $\cH_{r}$ (right panel) are plotted.  
  In the upper panels $\sigma=10^{-5}$~cm$^{-1}$, with results plotted for $t=0,0.3,1.0$~ms (red, blue, and black curves, respectively).  
  In the lower panels $\sigma=10^{-1}$~cm$^{-1}$, with results plotted for $t=0,0.4,1.0$~s (red, blue, and black curves, respectively).  
  Solid lines represent the analytical solution while open circles represent the element center value of the numerical solution.}
  \label{fig:diffusionProblem}
\end{figure}

@@

For the tests in this section we employ the tabulated opacities described in Section~\ref{sec:opacities}.  
We present two tests from the previous section (the homogeneous sphere and spherical diffusion problems), where the background is kept fixed.  
Then we present a relaxation problem involving neutrino-electron scattering (NES) with a homogeneous background.  
Finally, we present a deleptonization problem where the background density mimics that of a collapsed stellar core.  
The first deleptonization problem (Em/Ab) involves emission and absorption only, while the second deleptonization problem (Full) involves the full set of opacities described in Section~\ref{sec:opacities}.
